---
title: "GLM sheet 6"
output: html_notebook
---

In a psychology experiment, researchers asked participants to respond to various stimuli and
measured their reaction time. Participants were randomly assigned to one of three treatment
groups. Subjects in group 1 were required to respond as quickly as possible to any stimulus
that was presented. Subjects in group 2 were required to respond to a particular stimulus
while disregarding other types of stimuli. Finally, subjects in group 3 were required to respond
differently depending on the stimuli presented. The researcher felt that age may be a factor
in determining the reaction time, so she organized the subjects by age and obtained the
following data:

```{r}

 time <- c(0.384, 0.248, 0.191, 0.203, 0.331, 0.438, 0.494, 0.467, 0.302, 0.338, 0.495, 0.631, 0.485, 0.389, 0.629, 0.585, 0.782, 0.529, 0.586, 0.509, 0.364, 0.626, 0.858, 0.529, 0.52, 0.854, 0.7)
age <- gl(3,3,27)
group <- gl(3, 9, 27)
psychdata <- data.frame(cbind(time,age,group))
head(psychdata)
```

(a) Fit a two-way ANOVA model with interactions.

```{r}
fit <- glm(time ~ age*group, data = psychdata)

```

(b) Is there a significant interaction effect between group?
```{r}
summary(fit)
```
The p-value for the mixed effect, age:group, is 0.69 and so the coefficient is not significant.

(c) Draw interaction plots using the function interaction.plot and interpret the outcome;
see help("interaction.plot") for its usage. Does the output support the result of (b)?

```{r}
interaction.plot(age, group, response = time)
```
This suggests the mixed effect does have a small and inconsistent effect on the model fit, given that the mean is an increasing function of age for all groups. This largely supports my conclusion in (b).

(d) Refit the model without the interaction term. Has the significance of the main effects
changed? If so, why?

```{r}
fit2 <- glm(time ~ age + group, data = psychdata)
summary(fit2)
```
The two factors are now significant to the model. This is from removing the mixed effect, which has colinearity with both.

(e) What are the null and alternative hypotheses for the two main effects?

H0: Null Model vs. H1: Xi is in model

(f) Is there a significant difference in the mean reaction time between the three stimulus /
age groups?

```{r}
aggregate(time, list(age), mean, data=psychdata)

aggregate(time, list(group), mean, data=psychdata)
```
There is an evident increasing trend in the factor means.

(g) Make a residual plot. Do the assumptions of ANOVA appear to be valid?

```{r}
plot(fit2$residuals, psychdata$time)
```

There appears to be an increasing trend in the residuals, indicating a transformation of the response may be necessary. 