---
title: "LM Sheet 3 - part 2"
output: html_notebook
---

2. A hospital surgical unit was interested in predicting survival time in patients undergoing a
particular type of liver operation. From a random sample of 54 patients, information on the
patient's survival time, blood clotting score, prognostic index, enzyme function test score and
liver function test score were extracted. The data surgical.txt is available for download in
Blackboard.

```{r}
Surgical <- read.table(file="C:/Users/Dom/Documents/GitHub/Stats Work/Year 3/Linear Models/Surgical.txt", header=TRUE) #read in data in .txt format
head(Surgical) #view first six entries of Surgical
```


(a) Fit a multiple regression model using the survival time as the response variable and the
other four variables as explanatory variables.

```{r}
model <- lm(survival ~ blood.clotting + prognostic + enzyme + liver.function, data=Surgical) #fit full model
model
```


(b) Conduct an F-test for the overall fit of the regression model in (a). Comment on the
results.

```{r}
anova(model) #run F-test for full model
```


(c) Test each of the individual regression coefficients. Do the results indicate that any of
the explanatory variables can be removed from the model?


```{r}
summary(model)
```

  At the .05 significance level, the ANOVA table and the summary table for the full fitted model indicate that the intercept and the first three coefficients add significantly to the model's explanatory ability, while liver function does not. 

(d) Perform the forward, backward and backward-forward variable selection using the function
subset from the R package MASS. Compare the 'best' model selected from each
stepwise method, e.g., are all the methods return the same model? If not, why is it so?

```{r}
library(MASS) #load MASS package
null <- lm(survival ~ 1, data=Surgical) #specify null model
full <- lm(survival ~ ., data=Surgical) #specify full model
step(null, scope = list(lower=null, upper=full), direction = "forward") #forward selection
```
```{r}
step(full, data = Surgical, direction = "backward") #backward selection
```
```{r}
step(full, scope = list(lower=null), data=Surgical, direction = "both") #mixed selection
```

So each method respectively gave, as the chosen optimal model:
  lm(formula = survival ~ liver.function + prognostic + enzyme + 
    blood.clotting, data = Surgical)
  lm(formula = survival ~ blood.clotting + prognostic + enzyme, 
    data = Surgical)
  lm(formula = survival ~ blood.clotting + prognostic + enzyme, 
    data = Surgical)
    
These results occured due to the poor contribution liver.function makes to explanation.

(e) Using the function regsubsets from the package leaps to perform the subset selection
on the dataset and compare the results to the models obtained in (d). Comment on the
validity of the fitted model by plotting the residuals.

```{r}
library(leaps)
subset.fit <- regsubsets(survival ~., data=Surgical, nbest=10) #perform exhaustive model search
plot(subset.fit, scale = "adjr2") #plot results against adjusted R^2 measure
plot(subset.fit, scale = "Cp") #plot against Mallow's Cp
```

The adjusted R^2 measure suggests using the full model, while Mallow's Cp suggests omitting liver function. Both of these models were suggested by the methods employed in the previous question, implying validity in the results.

(f) Use the function boxcox from the R package MASS and suggest a good transform for the
response variable.

```{r}
bc <- boxcox(full)
bc$x[which.max(bc$y)]
```

Using lambda=0 is significant at the 0.05 level, so a log transform may be appropriate.

////////////////////////////////////////////

3. Produce the Lasso solution path on the Hitters data frame analysed in Section 7.3.1 of the
lecture notes using the function lars from the package of the same name, and compare it with
the Lasso solution path from glmnet. Also perform the 10-folds CV to select ?? that returns
the minimum CV error and compare the thus-selected ??blasso(??) to ??blasso(??
lasso
min ) reported in
Table 7.1. If they differ, explain why.

```{r}
library(ISLR)
head(Hitters)
new.hitters <- Hitters[!is.na(Hitters$Salary),] #omit missing data
y <- new.hitters[,19]
X0 <- new.hitters[, -19]
X0$League <- as.numeric(X0$League)-1 #covert entries to binary
X0$Division <- as.numeric(X0$Division)-1
X0$NewLeague <- as.numeric(X0$NewLeague)-1
X <- scale(X0) #centre and scale
```

```{r}
library(lars)
lars.model <-lars(X, y, type = "lasso") #use lars function to produce lasso path
plot.lars(lars.model) #plot paths
cv.lars(X, y, K=10) #compute 10-fold cross validation
```

