---
title: "LM Sheet 5"
output: html_notebook
---

2. The table below provides data on the number of grapevines affected by the fungus botrytis
cinerea, which is denote by yi out of mi grapevines treated with the strength, xi of a
particular chemical for groups i = 1, . . . , 7.

```{r}
m<- c(43, 41, 47, 66, 51, 57, 51)
y<- c(33, 29, 24, 20,16, 8, 11)
x<- c(1.0, 1.2, 1.5, 1.7, 2.0, 2.2, 2.4)
bcdata <- as.data.frame( cbind(m,y,x) )
bcdata
```

For each chemical strength, calculate the fitted values of the probabilities
that the grapevines are affected by the fungus and verify that the above maximum
likelihood estimates satisfy the likelihood equations.

```{r}
B0 <- 3.111; B1 <- -2.058
bcdata$fitted <- B0 + B1*bcdata$x
bcdata$fitted # show fitted values
bcdata$pi <- exp(bcdata$fitted)/ (1 + exp(bcdata$fitted))
bcdata$pi # show fitted probability estimates
```

```{r}
sum(bcdata$y)
sum(bcdata$m * bcdata$pi) #verify first equality

sum(bcdata$y * bcdata$x)
sum(bcdata$m * bcdata$pi * bcdata$x) #verify second
```

Calculate deviance
```{r}
A <-log(bcdata$y/bcdata$pi*bcdata$m)
C <-log( (1 - bcdata$y/bcdata$m)/(1 - bcdata$pi) )
sum(bcdata$y * A + (bcdata$m - bcdata$y)* C )
```
```{r}
pchisq(1096.791, 1, lower.tail = FALSE)
```

3.
The number of moths caught overnight in a trap is recorded on
20 randomly selected days. The average night time temperature is also recorded.

```{r}
y <- c(4, 6, 2, 7, 9, 5, 5, 8, 7, 1, 0, 1, 2, 4, 10, 11, 7, 7, 4, 4)
x <-  c(5, 10, 4, 6, 7, 3, 7, 4, 3, 3, 2, 4, 6, 8, 13, 11, 6, 3, 3, 3)
moth <- as.data.frame(cbind(x,y))
moth
```

We would like to model the number of moths using a GLM with average night time temperature
as a covariate.

```{r}
model <- glm(y ~ x, data = moth, family = "poisson")
model
```

Find fisher information matrix

```{r}
sum(-exp(1.0270 + 0.1028*moth$x))
sum(-(moth$x) * exp(1.0270 + 0.1028*moth$x))
sum(-(moth$x^2) *exp(1.0270 + 0.1028*moth$x))

```
Predict for x = 6.5
```{r}
predict.glm(model, newdata = data.frame(x=6.5), type = "response", se.fit = TRUE)
```

