---
title: "GLM sheet 7"
output: html_notebook
---

```{r}
#clinic <- c("A","A","B","B")
#care <- c("Low","High","Low","High")
#survived <- c(176,293,197,23)
#died <- c(3,4,17,2)
#counts <- survived + died
#antenatal.data <- data.frame(cbind(clinic,care,survived,died,Counts))
antenatal.data <- read.table("C:/Users/Dom/Downloads/antenatal.data", header=TRUE)
attach(antenatal.data)
fit1 <- glm(counts~clinic+care+survival, family="poisson")
summary(fit1)
```

The first task is to investigate whether survival is dependent on the clinic. Write down
the model expression in which the appropriate interaction is added to the additive model.
Use anova to test for an association between survival and clinic. Based on the analysis
of deviance table, decide whether to include the interaction term in the model.

```{r}
fit2 <- glm(counts~care+survival:clinic, family="poisson", data=antenatal.data)
summary(fit2)
anova(fit1,fit2)
```

The interaction term is shown by the deviance table to be highly significant, so this should be included.

Add a term to the model in (d) to investigate a possible association between level of care
and clinic. Write down the new model and decide whether to include the extra term in
the model based on an analysis of deviance.

```{r}
fit3 <- glm(counts~ survival:clinic + care:clinic, family="poisson", data=antenatal.data)
summary(fit3)
```

