---
title: "Model selection with the Housing dataset"
output: html_notebook
---

The dataset Housing from the R package Ecdat contains sales prices of houses in
the city of Windsor, Canada in 1987, with n = 546. Eleven variables were observed
about various aspects about the houses, including the size of a property, the number of
bedrooms/bathrooms and so forth; details of the explanatory variables can be found
in help(Housing).


```{r}
#install.packages("Ecdat")
library(Ecdat) #access Ecdat
data(Housing) #access Housing
head(Housing) #view first 6 entries of Housing
```


When the number of candidate variables q is small, it may be possible to fit all 2^q, q=11, submodels and select the one that meets your chosen criterion. The function regsubsets
in the package leaps implements regression subset selection and allows one to view
the ranked models according to different scoring criteria by plotting the results of
regsubsets.

```{r}
#install.packages("leaps")
library(leaps) 
subset.fit <- regsubsets(price ~ ., data = Housing, nbest = 10) #fit all possible models for price
#nbest specifies the number of models to be shown for each number of parameters -e.g the amount of models to consider for p=3
par(mfrow = c(1, 2)) #set graphic scale
plot(subset.fit, scale = 'adjr2') #plot results against adjusted R^2
plot(subset.fit, scale = 'Cp') #plot results against Mallow's Cp

```

According to both criteria, the model containing lotsize, bathrms, stories, fullbase, gashw, airco, garagepl and prefarea are
selected.

The exhaustive search performed by regsubsets becomes quickly impossible with
increasing q. The step function in the package MASS automates the process of backward,
forward, or backward-forward selection based on AIC. To perform forward
selection, we need to specify a starting model and the range of models which we want
to examine in the search. The following code tells R to start with the null model and
search through models lying in the range between the null and full models using the
forward selection algorithm.

```{r}
library(MASS)
null <- lm(price ~ 1, data = Housing) #specify null model
full <- lm(price ~ ., data = Housing) #specify full model
step(null, scope = list(lower = null, upper = full), direction = "forward") #conduct forward selection

```

According to this procedure, the best model is the one that includes all eleven variables.
We can perform backward elimination or stepwise regression on the same data
set using the commands:

```{r}
step(full, data=Housing, direction = "backward") #conduct backward selection
step(full, scope=list(lower = null), data=Housing, direction = "both") #conduct forward-backward selection
```

These give equivalent results to forward selection.