---
title: "R Notebook"
output: html_notebook
---

Problem 3 (ARCH models)
Let p1, . . . , p101 be the daily closing prices of the S&P 500 index between 18/10/2018 and
15/03/2019, and let xt = log(pt+1/pt), for t = 1, . . . , 100, be the corresponding daily logreturns. In this problem we use an ARCH (Generalized Auto-Regressive Conditionally
Heteroscedastic) model to model the observation x := (x1, . . . , x100). An ARCH(q)
model, which is a popular model for the log-returns of a financial asset, defined by X_t.

. Let fq(x|??q) be the likelihood function of
the model. We make the standard assumption that Pq
i=1 ??i < 1, which implies that
the stochastic process (Xt)t???1 dened in (1) is stationary. To incorporate this prior
information we take as prior distribution for (??1, . . . , ??q) the uniform distribution on the
set
{y ??? R
q
>0
:
X
q
i=1
yi < 1
	
}
We let the Gamma(1, 1/2) distribution be the prior distribution for ?? and use the notation
??q(??q) for the so-dened prior distribution on ??q. It can be shown that, for i = 1, . . . , q,
the marginal distribution of ??i under ??q(??) is the Beta(1,q) distribution.
The data1 are in the le SP500.txt and can be loaded in R as follows (assuming that
your .R le is in the same folder as the DataSheet5 folder):

```{r}
# Load SP500 data
p<-read.table('DataSheet5/ARCH/SP500.txt')
# Compute log-returns
x<-diff(log(p[,5]))

```

1. Plot all the marginal distributions of ??q(??q) for q = 1, 2. Is the prior distribution
for ??1 non-informative in the sense of Laplace for q ??? {1, 2}?

```{r}
xset<-seq(0, 1, 0.01)
plot(xset, dbeta(xset, 1, 1 ), type = "l")
plot(xset, dbeta(xset, 1, 2 ), type = "l")
```

q = 1 has a non-informative prior, but q = 2 places more density at lower values of x.


2. Consider first an ARCH(1) model.
a) Implement in R a Metropolis-Hastings algorithm having
??1(??1|x) ??? f1(x|??1)??1(??1)
as invariant distribution, starting value ??1,0 = (0.01, 0.1) and where the proposal distribution Q(??1, d~??1) is the N2(??1, ??1) distribution. Propose a matrix
??1, a burn-in period B and a length T for the simulated trajectory (burn-in
period non-included). Justify your choices.
```{r}
# Load the package MASS that can be used to sample from
# multivariate normal distributions
library(MASS)
# Density of the distribution we want to sample from (in log)
density_target<-function(y){
return(dnorm(y[1],y[2],1,log=T))
}
# Function that samples from the proposal distribution
simulate_proposal<-function(y, Sigma){
return(mvrnorm(1,y,Sigma))
}
# Parameters to choose
T<-150000 #length of the trajectory
y0<-c(0.01,0.1) #starting value
c<-2.7 #THE CONSTANT TO CALIBRATE
Sigma <- c*matrix(c(1,0.25,0.25,1),2,byrow=T) #covariance matrix of Q
# To store the results
chain<-matrix(0,T,2)
acceptance_rate<-0
#Run the Metropolis-Hastings algorithm
chain[1,]<-y0

mu0<-density_target(chain[1,])
for(t in 2:T){
yt<-simulate_proposal(chain[t-1,],Sigma)
mu1<-density_target(yt)
if(log(runif(1))<=(mu1-mu0)){
chain[t,]<-yt
mu0<-mu1
acceptance_rate<-acceptance_rate+1
}else{
chain[t,]<-chain[t-1,]
}
}
acceptance_rate<-acceptance_rate/T
```

```{r}
plot(1:T, chain[,1], xlab='t', ylab=expression(y[1]), type='l')
plot(1:T, chain[,2], xlab='t', ylab=expression(y[2]), type='l')
```

b) Plot the estimated marginal distributions of ??1(??1|x) and give a 99% condence interval for each of the corresponding posterior means.

```{r}
B<- 1500
sample <- as.mcmc(chain[B:T,])
plot(sample[,1], )
plot(sample[,2])

```





3. Consider now an ARCH(2) model.
a) Implement in R a Metropolis-Hastings algorithm having
??2(??2|x) ??? f2(x|??2)??2(??2)
as invariant distribution, starting value ??2,0 = (0.01, 0.1, 0.1) and where the
proposal distribution Q(??2, d~??2) is the N3(??, ??2) distribution, with ??2 the
matrix given in the le Sigma_2.txt:
```{r}
Sigma<-as.matrix(read.table('C:/Users/Dom/Documents/GitHub/Statistics/Year 3/Bayesian Analysis/DataSheet5/ARCH/Sigma_2.txt'))
```

Propose a burn-in period B and a length T for the simulated trajectory (burnin period non-included). Justify your choices.

```{r}
# Density of the distribution we want to sample from (in log)
density_target<-function(y){
return(dnorm(y[1],y[2],y[3],log=T))
}
# Function that samples from the proposal distribution
simulate_proposal<-function(y, Sigma){
return(mvrnorm(1,y,Sigma))
}
# Parameters to choose
T<-50000 #length of the trajectory
y0<-c(0.01,0.1,0.1) #starting value
#c<-0.01 #THE CONSTANT TO CALIBRATE
#Sigma<- matrix(c(1,0.6,0.6,1),2,byrow=T) #covariance matrix of Q
# To store the results
chain<-matrix(0,T,3)
acceptance_rate<-0
#Run the Metropolis-Hastings algorithm
chain[1,,]<-y0

mu0<-density_target(chain[1,,])
for(t in 2:T){
yt<-simulate_proposal(chain[t-1,,],Sigma)
mu1<-density_target(yt)
if(log(runif(1))<=(mu1-mu0)){
chain[t,]<-yt
mu0<-mu1
acceptance_rate<-acceptance_rate+1
}else{
chain[t,,]<-chain[t-1,,]
}
}
acceptance_rate<-acceptance_rate/T
```


b) Plot the estimated marginal distributions of ??2(??2|x) and give a 99% condence interval for each of the corresponding posterior means.




4. Consider now the test, in the ARCH(2) model, of the hypothesis H0 : ??2 ??? 0.1
against the alternative H1 : ??2 > 0.1.
a) Provide an estimate and a 99% condence interval for ??2

{??2 : ??2 ??? 0.1}|x).
Do you accept H0 under the a0 ??? a1 loss function when a0 = a1?
b) Compute an estimate of B
??2
01 (x), the Bayes factor for the test we are considering.
c) Comment the results obtained in parts 4.a) and 4.b).





5. Based on the results of part 4, choose between the ARCH(1) model and the
ARCH(2) model