---
title: "Bayesian Modelling  Problem Sheet 6"
output: html_notebook
---

Radon is a natural radioactive gas that is usually found in rocks and soils. Radon
gas, which is for most people the single largest source of radiation, is known to be a
source of lung cancer when breathed in high concentration. In particular, according to
the United States Environmental Protection Agency (EPA), radon is the most frequent
cause of lung cancer among non-smokers. Due to local dierences in geology, the level of
the radon gas hazard diers from location to location, and it is therefore important to
identify areas with high radon exposures to help home-owners to make decisions about
measuring or remediating the radon in their house. More information about radon and
its health eects can be found here: www.radon.com/radon_facts. 
The dataset radon.txt contains radon measurements taken by the EPA in 917 randomly
chosen houses located in 84 (out of 87) counties in the state of Minnesota. 


```{r}
radon <- read.table('DataSheet6/radon.txt', sep='')
set.seed(333)
library(MASS)
library(rjags)
library(coda)
load.module('glm')

```

The first column of the dataset contains the 917 measurements of radon (in log) while
the second column indicates if a measurement was taken in a basement (value 1) or in a
first floor (value 0). This information is important to understand the level of radon in a
given house since, as the radon comes from the soil, it enters more easily in a house when
it is built into the ground. Columns three and four provide respectively the county name
and the county index of the sampled house and the last column of the dataset gives a
measurement of soil uranium (the ultimate source of radon) taken at the county level.

The objective of this assignment is to use a Bayesian approach to solve the following
two problems:

P1 Predict the log-radon level of a house with a basement for each of the 84 counties
in the dataset.
P2 Predict the log-radon level of a house with no basement and of a house with a
basement in the county of Cook, for which no measurements have been taken.

2. Estimate your model. In particular:
a) (15 marks) Implement your model in JAGS.


```{r}
#write model to radon.bug
cat('
model{  
for(j in 1:M){
theta[j] ~ dnorm(alpha*uu[j], sigma2^(-2))
}
for(i in 1:n){
x[i] ~ dnorm( (theta[county.index[i]]) + beta*B[i], sigma^(-2))
}

alpha ~ dnorm(0,0.0001)
beta ~ dnorm(0,0.001)
sigma ~ dunif(0,100)
sigma2 ~ dunif(0,10000)
}', file='radon.bug')

```
```{r}
#specify determined parts
radon_data<-list(n=nrow(radon), M=max(radon$county.index), x=radon$log.radon, county.index = radon$county.index,
B=radon$floor, uu = unique(radon$uu))
#specify prior points
radon_inits <- list(alpha=0,beta=0,sigma=1,sigma2=1)
```
```{r}
#construct model in jags
radon_mu<-jags.model('radon.bug',inits = radon_inits,data=radon_data)
#verify we are sampling correct quantities, and using correct samplers
list.samplers(radon_mu)
```



b) (15 marks) Simulate a trajectory of length T and justify your choice for T
```{r}
#set burn-in length
Burn <- 2000
#set trajectory length
Tr <- 10000
#update chain by burn-in length
update(radon_mu, n.iter = Burn)
#sample from distribution with correct quantities
sample<-coda.samples(radon_mu, variable.names = c("alpha","theta","beta","sigma","sigma2"), n.iter=Tr)
```

```{r}
#plot autocorrelations for parameters to verify convergence
autocorr.plot(sample[,1:7], 100)
#These have rapidly-declining autocorrelation, indicating the chain has converged for all parameters
```


```{r}
traceplot(sample[,1:7])
#These appear as white noise, centred on the parameter estimate, indicating the chain has converged to a stationary distribution
summary(sample)$statistics
#All the standard deviations are small, supporting convergence 
```


3. Provide an answer to the above problem P1. In particular,
a) (5 marks) For each county in the dataset, compute the average log-radon
level of a house with a basement. Make a plot with these estimates on the
y-axis and the number of observations per county on the x-axis. Add in the
plot an horizontal line that represents the average log-radon level of a house
with a basement for Minnesota.
```{r}
#select houses with basements
radon_bm <- radon[radon$floor==1,]
#find average and counts by county 
county_means <- rep(0, 84)
county_count <- rep(0, 84)
for(i in 1:84){
  county_means[i] <- mean(radon_bm$log.radon[radon_bm$county.index == i])
  county_count[i] <- nrow(radon_bm[radon_bm$county.index == i,])
}
```
```{r}
#plot means against counts
plot(county_count,county_means)
#find overall mean and overlay as line
state_mean <- mean(radon_bm$log.radon)
abline(a=state_mean, b=0)
```

b) (15 marks) Using your model, predict for each county in the dataset the
log-radon level of a house with a basement. Explain how these predictions
have been computed from your model and make a plot with these estimates
on the y-axis and the number of observations per county on the x-axis. Add
in the plot an horizontal line that represents the average log-radon level of a
house with a basement for Minnesota.
```{r}
#select estimates for parameters as a new vector
pars <- summary(sample)$statistics[,1]
#predict log radon for each county, assuming house has a basement, using model equation
preds <- rep(0, 84)
for(i in 1:84){
  preds[i] <-  pars[1]*pars[i+4] + pars[2]
}
```
```{r}
#plot predictions against counts by county
plot(county_count, preds)
abline(a=state_mean, b=0)
```


c) (5 marks) Compare the `naive' estimates obtained in part 3.a) with the model
based estimates obtained in part 3.b). Comment your results.
```{r}
#plot estimates against each other
plot(county_means, preds)
#find mean square error of both estimates
mean((county_means-preds)^2)
```



4. (15 marks) Provide an answer to the above problem P2, that is use your model
to predict the log-radon level of a house with no basement and of a house with a
basement located in Cook, given that the soil uranium measurement for this county
is equal to -0.504995982525783. Explain how your prediction has been computed
from your model.
```{r}
u_cook <- -0.504995982525783
#Predict the log-radon level of a house with no basement in the county of Cook
pars[1]*u_cook 

#Predict the log-radon level of a house with a basement in the county of Cook
pars[1]*u_cook + pars[2]
```

