---
title: "Classification of Iris Flowers"
output:
  html_document:
    df_print: paged
---
///
Credit to machinelearningmastery.com/machine-learning-in-r-step-by-step/
///

As an introduction to machine learning, I will construct a multi-variable regression model with the intent of classifying a dataset of iris flowers. This dataset contains 150 observations of flowers, and 5 characteristics associated with each - 4 numeric measurements, and the species.

```{r initialise}
#load the caret package (http://topepo.github.io/caret/index.html)
library(lattice)
library(ggplot2)
library(caret)

# attach the iris dataset to the environment
data(iris)
# rename the dataset
dataset <- iris

```

I will need a dataset for validation of the model, and so I will partition the entire dataset into two parts, 80% for training and 20% for validation.

```{r validation set}
# create a list of 80% of the rows in the original dataset we can use for training
validation_index <- createDataPartition(dataset$Species, p=0.80, list=FALSE)
# select 20% of the data for validation
validation <- dataset[-validation_index,]
# use the remaining 80% of data to training and testing the models
dataset <- dataset[validation_index,]
```
Let's explore the data we have.

```{r explore}
# dimensions of dataset
dim(dataset)

# list types for each attribute
sapply(dataset, class)

# take a peek at the first 5 rows of the data
head(dataset)

# list the levels for the class - i.e, the iris species
levels(dataset$Species)

# summarize the class distribution
percentage <- prop.table(table(dataset$Species)) * 100
cbind(freq=table(dataset$Species), percentage=percentage)

# summarize attribute distributions
summary(dataset)
```

Now, let's visualise the data and get a better look at it.

I'll start with univariate plots, to better understand the properties of each variable.

```{r visualise 1}
# split input and output - the measurements define x, and the classes, y
x <- dataset[,1:4]
y <- dataset[,5]

# boxplot for each attribute on one image
par(mfrow=c(1,4))
  for(i in 1:4) {
  boxplot(x[,i], main=names(iris)[i])
}

# barplot for class breakdown
plot(y)
```
This tells me about the measurement distributions, and that the observations are equal in number.

///

Multivariate plots can tell me about the relationships between the different variables.

```{r visualise 2}
# scatterplot matrix - this creates a chart of each variable plotted against every other, allowing trend recognition
featurePlot(x=x, y=y, plot="ellipse")

# box and whisker plots for each attribute - as in the above boxplots, this allows distribution comparison, but also can identify clear differences between classes
featurePlot(x=x, y=y, plot="box")

# density plots for each attribute by class value - from this, I can see the distribution of each attribute in a mooth manner
scales <- list(x=list(relation="free"), y=list(relation="free"))
featurePlot(x=x, y=y, plot="density", scales=scales)
#These look somewhat Gaussian!
```

Now that I have a deeper understanding of the data properties, it's time to set up some ML algorithms for predicting the class of each flower from its attributes.

I'll split the dataset into 10: 9 for training, 1 for for testing. I will repeat each process 3 times, for accuracy.

```{r test harness}
# Run algorithms using 10-fold cross validation
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
```

I'll use the 5 suggested algorithms for my dataset: LDA, CART, kNN, SVM, RF


```{r build models}
# a) linear algorithms
set.seed(7)
fit.lda <- train(Species~., data=dataset, method="lda", metric=metric, trControl=control)
# b) nonlinear algorithms
# CART
set.seed(7)
fit.cart <- train(Species~., data=dataset, method="rpart", metric=metric, trControl=control)
# kNN
set.seed(7)
fit.knn <- train(Species~., data=dataset, method="knn", metric=metric, trControl=control)
# c) advanced algorithms
# SVM
set.seed(7)
fit.svm <- train(Species~., data=dataset, method="svmRadial", metric=metric, trControl=control)
# Random Forest
set.seed(7)
fit.rf <- train(Species~., data=dataset, method="rf", metric=metric, trControl=control)
```

And then compare them in terms of accuracy

```{r compare accuracy}
#summarise accuracy of models
results <- resamples(list(lda=fit.lda, cart=fit.cart, knn=fit.knn, svm=fit.svm, rf=fit.rf))
summary(results)

# compare accuracy of models
dotplot(results)
```
Hence, the LDA is the most accurate. I'll zoom in on just this model.

```{r summarise LDA}
# summarise Best Model
print(fit.lda)
```

Using the LDA model, I want to get an idea of the accuracy of the model on the validation set, to quantify how well my choices have performed.

```{r predict}
# estimate skill of LDA on the validation dataset
predictions <- predict(fit.lda, validation)
confusionMatrix(predictions, validation$Species)
```

This is 100% accurate! This meets the expected criterion of 97 +- 4% for a model, suggesting this is a suitable choice for our dataset. Using this model, I can classify the species of any iris, based on the 4 measurements needed, with perfect accuracy.

///

To conclude, I have followed a tutorial guiding me through the principles of a machine learning project. I have explored and visualised a dataset, created 5 ML models, evaluated their accuracy, and verified that my chosen model is suitable. I now have the skills and the framework to perform further analyses on other data, using other models.


