---
title: "TSA Lab 2"
output: html_notebook
---

1. Simulate observations from a purely random process


```{r}
z = rnorm(n=300, mean=0, sd=1)
ts.plot(z, main="A purely random process")

```

```{r}
# Define our own version of the autocorrelation
rk=function(x,k) {
n = length(x)
x1 = x[1:(n-k)]
x2 = x[(1+k):n]
meanx = mean(x)
r = sum((x1-meanx)*(x2-meanx))/sum((x-meanx)^2)
return(r)
}
# Try out our version
rk(z,1)
rk(z,2)
# R has its own function to compute it: acf(), let's try it...
z.acf=acf(z,plot=T)
z.acf
```

2. Simulate observations from a random walk
```{r}
# Show the help for the cumsum() function
?cumsum
# Now simulate a random walk...
x <- cumsum(rnorm(n=300, mean=0, sd=1))
# ... and now plot the simulated random walk...
ts.plot(x,main="Random walk")
```

3. Simulate observations from a moving average process

```{r}
# Simulate the MA(2) process Xt = Zt ??? 1.3Zt???1 + 0.4Zt???2
ma.sim=arima.sim(model=list(ma=c(-1.3,0.4)), n=300)
# Plot the time series and the correlogram of this simulated series
par(mfrow=c(2,1))
ts.plot(ma.sim, main="MA(2)")
ma.acf = acf(ma.sim, type="correlation", plot=T)
ma.acf

```

4. Simulate observations from an autoregressive process

```{r}
# Simulate N = 300 observations from an AR(1) process defined as Xt = 0.9Xt???1 + Zt
ar1p.sim = arima.sim(model = list(ar=c(0.9)), n = 300)
par(mfrow = c(2,1))
ts.plot(ar1p.sim, main = "AR(1) with coefficient 0.9")
# Let's look at this simulated data's sample acf, and see if it matches the model value.
ar1p.acf = acf(ar1p.sim, plot = T)
ar1p.acf
# Calculate sample partial autocorrelation coefficient
ar1p.pacf = pacf(ar1p.sim, plot = T)
ar1p.pacf

```

Exercises

1.Simulate N = 300 observations from Xt = ???0.9Xt???1 + Zt, where {Zt} is a purely randomprocess with mean 0 and variance 1. Plot the simulated time series and its correlogram.

```{r}
ar1p.sim = arima.sim(model = list(ar=c(-0.9)), n = 300)
par(mfrow = c(2,1))
ts.plot(ar1p.sim, main = "AR(1) with coefficient -0.9")

ar1p.acf = acf(ar1p.sim, plot = T)
ar1p.acf
```


2. Simulate N = 300 observations from an AR(2) process defined as Xt = 0.9Xt???1 ???0.2Xt???2 +Zt,where {Zt} is a purely random process with mean 0 and variance 1. Plot the simulated timeseries, its correlogram and partial autocorrelation coefficients.

```{r}
ar2p.sim = arima.sim(model = list(ar=c(0.9, -0.2)), n = 300)
par(mfrow = c(3,1))
ts.plot(ar1p.sim, main = "AR(2) with coefficients 0.9, -0.2")

ar2p.acf = acf(ar2p.sim, plot = T)
ar2p.acf

ar2p.pacf = pacf(ar2p.sim, plot = T)
ar2p.pacf
```

