---
title: "TSA lab 4"
output: html_notebook
---



```{r}
load(url("https://people.maths.bris.ac.uk/???magpn/TSA/TSA.RData")) #access data
```

1 Fitting an ARIMA model

This is a continuation of Example 2 in Lab One. The land-air temperature anomalies (from 1961-
1990 average), measured in degrees centigrade, for the years 1900-1997, are stored in the data set
gtemp. We follow the flowchart to identify whether an ARIMA model is appropriate for this data
set.


(a) Plot the time series. Does it look stationary in mean? If not, difference the data using the
R function diff() and put it into another vector called dgtemp. Plot the differenced data.
Does it appear stationary? If so, then set the order of differencing d = 1, otherwise difference
again and increase d accordingly. (Actually, d = 1 looks reasonable).

```{r}
ts.plot(gtemp) #plot gtemp data

#this doesn't look stationary

dgtemp <- diff(gtemp) #difference gtemp
ts.plot(dgtemp) # plot differenced data

#this does look stationary
```

(b) Plot the correlogram of the differenced data using the acf() function. Are any of the correlation
coefficients much larger than the tolerance limits?

```{r}
acf(dgtemp) #plot autocorrellation of differenced gtemp

```

None of the lags are significantly large, other than k=0.

(c) Plot the partial correlogram of the differenced data. Are any of the partial autocorrelation
coefficients much larger than the tolerance limits?

```{r}
pacf(dgtemp) #plot partial autocorrellation of differenced gtemp

```

Lags k=2 and k=3 are significantly larger than the tolerance interval.

(d) Suggest a suitable model for the difference data and write down the fitted model. Hint: The
fit can be obtained with the arima() function.

I have that d=1 since I have differenced the series once to obtain a stationary sequence. 
The PACF indicates I should use p=3 for the AR parameter, as the last significant time step with correllation not accounted for by the ACF is at step 3.
The ACF implies I should use q=0, as there are no significant correlations with previous time steps.
Hence, I use the model ARIMA(3,1,0) for the gtemp series.

```{r}
arima(gtemp, order = c(3,1,0))
```

So I have the model dXt = -0.3351Xt-1 -0.3245Xt-2 -0.3388Xt-3 +Zt

//////////////////////////////////

2 Exponential smoothing

(a) Implement the es() function now in R (type in the above definition of es() just as it appears.
You can type RETURN wherever any space occurs to split the typing over more than one
line).

```{r}
es <- function(x, alpha){sum(rev(x)*alpha*(1-alpha)^(0:(length(x)-1)))} #define es function
```


(b) Now create a new data vector from the Lynx data by copying all the observations in the Lynx
data set apart from the last one by using the command
lynx2 <- lynx[1:113]
The last value of the lynx data set is lynx[length(lynx)] which is l114 = 3396.

```{r}
lynx2 <- lynx[1:113] #define 1-step lynx vector
```


(c) Now use exponential smoothing on lynx2 with varying values of ??. For example, es(lynx2,0.2),
try some others. This can be done by the following R commands.
forecasts=NULL
alphas=seq(from=0.1, to=1, by=0.05)
for(i in alphas) forecasts=c(forecasts, es(lynx2,i))
plot(alphas, forecasts)
The plot suggests that the best value of ?? = 1 as this is closest to the true value l114 = 3396.
However, the above procedure is not very good for the lynx data set because of the seasonality
in the data.

```{r}
forecasts=NULL
alphas=seq(from=0.1, to=1, by=0.05) #define incremental vecotr from 0.1 to 1
for(i in alphas) forecasts=c(forecasts, es(lynx2,i)) #apply es to each alpha value
plot(alphas, forecasts) #plot alphas against exponential smooth

```


(d) To try and correct for the obvious cyclic behaviour (every 10 years?) in the series, create a
new series from the lynx set by differencing the data with a lag of 10. In other words, if the
lynx series is represented by lt form
dt = lt ??? lt???10 for t = 11, . . . , 114.
In R you can do this by
lynx10 <- diff(lynx, lag=10)
Plot the new differenced time series (use ts.plot). This looks less seasonal.

```{r}
lynx10 <- diff(lynx, lag=10) #difference lynx at lag k=10
ts.plot(lynx10) #plot differenced series
```


(e) Suppose we are interested in finding a forecast for LN+1. Rewrite the formula in part (d) to
get
Lt = Dt + Lt???10.
We have capitalized letters because we are dealing with the idea of what the model is behind
the process. Then stepping one time step into the future gives
Lt+1 = Dt+1 + Lt???9.
When we are trying to forecast LN+1 we just set t = N into the equation, i.e., LN+1 =
DN+1 + LN???9.



(f) Use exponential smoothing on the Dt to get a "sensible" estimate for DN+1. Then we may
use this estimate and the actual value lN???9 to obtain an estimate of LN+1.
Suppose that we have lynx2, i.e, l1, l2, . . . , l113, and we are trying to forecast l114 = 3396 (so
pretend we don't know l114 for the time being). Then using the diff command it is possible
to compute
dt = lt ??? lt???10 for t = 11, . . . , 113.
The formula for the estimate of L114 is
L114 = D114 + l104
(we are pretending l114 is unknown at this stage). We know l104 from the data and we can
estimate D114 from exponential smoothing of the dt
, for t = 11, . . . , 113. We can perform the
estimation for L114 by using the R commands

```{r}
lynx10.2=diff(lynx2,lag=10) #define copy of lynx10
forecasts=NULL
alphas=seq(from=0.1, to=1, by=0.01)
for(i in alphas) forecasts=c(forecasts,lynx[104]+es(lynx10.2,alpha=i))
plot(alphas, forecasts)
abline(h=lynx[114])
locator(1) 
```

The plot from these commands is shown in Figure 1. The last line in the commands plots a
horizontal line at the value of l114. You can see that the closest forecast is around the value of 0.33. In fact, you can find out by typing locator(1)
which allows you to click on the plot with the mouse, after clicking R returns the coordinates
on the plot where you clicked. Therefore, our best choice of alpha is 0.33. You can now
forecast L114 by typing

```{r}
lynx[104]+es(lynx10.2,alpha=0.33)
```

and you should get ^l(113, 1) ??? 3393, which is better than the previous estimate.

You should also try the following R commands to forecast by using the AR model which we fitted
in Lab Three to the log of the lynx data. This method is essentially Box-Jenkins forecasting.

```{r}
llynx=log(lynx)
llynx2=llynx[1:113]
fit=arima(llynx2,order=c(11,0,0))
predict(fit,n.ahead=1)
```

Exercise. Recife Dataset
(a) Create a new data vector from the recife data by copying all the observations apart from
the last 12 observations by using the command
recife2=recife[1:108]
We are pretending the last 12 observations are unknown. We'll fit a model to recife2 and
find forecasts for up to 12-steps ahead.

```{r}
recife2=recife[1:108] #define shorter recife vector
ts.plot(recife2)
```


(b) Estimate the seasonal effects of recife2. Then work with the deseasonalized data. Follow
the ARIMA fitting flowchart in your notes to suggest a suitable model for the deseasonalized
data and write down the estimated model.

```{r}
acf(recife2) # plot acf of recife2
#this indicates a seasonality period of 12
recife2.12 <- diff(recife2, lag = 12) #define vector of differenced series at lag 12

ts.plot(recife2.12) #plot deseasonalised series
#this appears stationary, so d=0

acf(recife2.12) #plot acf of recife2.12
#These are significant to lag 3, so use q=3 (lag 12 is significant, but I know the series holds a seasonal property so I can ignore this)
#This does appear to cut off, though I'll investigate the pacf to determine whether a non-zero parameter p is needed

pacf(recife2.12) #plot pacf
#this cuts off at p=1, so I'll use this parameter in the model.

```

I now have reason to believe the ARMA(1,3) model is suitable.

```{r}
fit <- arima(recife2, order = c(1,0,3), seasonal = list(order = c(0, 0, 0), period = 12)) #fit ARMA(1,3), seasonal model to the series
fit
```

So I have the model Xt = 0.4383Xt-1 +Zt  +0.5403Zt-1  +0.5207Zt-2  +0.4292Zt-3    +25.7579

(c) Find forecasts for up to 12-steps ahead. (Don't forget to add back the seasonal effects.)

```{r}
predict(fit, n.ahead = 12)
```

