---
title: "TSA Sheet 6"
output: html_notebook
---
3.
Create a purely random normally distributed time series realization with mean 0 and variance 1
```{r}
long <- rnorm(8192)

```

Plot the data with
```{r}
ts.plot(long)

```

Compute and plot the raw periodogram
```{r}
rawp <- spec.pgram(long, taper = 0, detrend = FALSE,
plot = FALSE)
plot(rawp)

```

The periodogram plotted in the previous picture is on a log scale, and does not look much like a
set of ??
2
random variables. We can turn this off using the following command
```{r}
plot(rawp, log="no")
```

In any case, both spectra look much too rough. R can smooth the spectra by using a technique
called modified Daniell smoothing. In R the degree of smoothing is controlled by the spans
argument to spec.pgram(). So try
```{r}
smop <- spec.pgram(long, spans=c(9,7), taper = 0,
detrend = FALSE, plot = FALSE)
plot(smop, ylim=c(1e-3, 10))

```
4.
Compute the log of the lynx data using
```{r}
llynx <- log(lynx)
ts.plot(llynx)
```

Guess what the period of the sinusoidal variation is - 
This appears to be around 10 years.

Compute the raw periodogram of the data (call it rawpl) and plot it. What do you see?
```{r}
rawpl <- spec.pgram(llynx, taper = 0, detrend = FALSE)
```

There should have been a peak at about 0.1. However, the raw periodogram is quite noisy so try
the following two methods to smooth the periodogram:
i. Use the spec.pgram() function with the spans=c(9,7) argument (as in the previous
question).
```{r}
spec.pgram(llynx, taper = 0, detrend = FALSE, spans = c(9,7))
```

```{r}
smopl2 <- spec.pgram(llynx, kernel=kernel("daniell", c(1,2)),
taper = 0, detrend = FALSE, plot = FALSE)
plot(smopl2)

```

To find out more accurately where the largest peak is. Type > locator(1)
after you have plotted the smoothed spectrum in either case and click the mouse at the maximum
value on the graph. The x and y coordinates of where you clicked will be returned allowing you
to identify the frequency where the periodogram maxed out

```{r}
smopl2 <- spec.pgram(llynx, kernel=kernel("daniell", c(1,2)),
taper = 0, detrend = FALSE, plot = FALSE)
plot(smopl2)
locator(1)
```

Using the spans smoothing method the maximum power occurs at around a frequency of 0.101.
The equivalent peak in the kernel method is at about 0.103. Since the original data are measured
in years this corresponds to approximately 9.6-9.9 years. Does this correspond to your
original guess as to the period of the sinusoidal frequency of the lynx data? Note that the answer
of anything between 9.6 and 9.9 years cannot be obtained directly from the time series or methods
based in time like the autocorrelation function (i.e. we guessed somewhere between 9 and
10 years) - 
This does correspond to my observational estimate.

