---
title: "Assignment 3"
output: html_notebook
---
Michelson and Morley performed a classic experiment which tested the hypothesis that the earth
travels through a luminiferous aether. As part of this experiment, they made many measurements
of the speed of light.


```{r}
#load the data
library(datasets)
data(morley)
lightspeed <- morley$Speed
```

We test the hypothesis that the observations in lightspeed come from independent identical normal
distributions, using Pearson’s goodness of fit test and discrete quantisations.

```{r}
#quantise
n <- length(lightspeed)
breaks <- c(-Inf, seq(from = 745, to = 945, by = 50), Inf)
Obs <- table(cut(lightspeed, breaks))
Obs

```
We test whether these results are consistent with a Multinomial(100, p(mu, sigma^2) ),

So H0: q is in { p(mu,sigma^2) } 
against H1: q is not in { p(mu,sigma^2) }

The ML estimate for (mu, sigma^2) is (sample mean, sample variance)

```{r}
#obtain maximum likelihood estimate for mean and variance
Meanml <- mean(lightspeed)
Varml <- var(lightspeed)

cat("Sample mean = ",  Meanml); 
cat("Sample var = ",  Varml); 
#create the vector q for testing
q = rep(0, 6)
q[1] = pnorm(745 ,Meanml, sqrt(Varml))
q[2] = pnorm(795 ,Meanml, sqrt(Varml)) - pnorm(745 ,Meanml, sqrt(Varml))
q[3] = pnorm(845 ,Meanml, sqrt(Varml)) - pnorm(795 ,Meanml, sqrt(Varml))
q[4] = pnorm(895 ,Meanml, sqrt(Varml)) - pnorm(845 ,Meanml, sqrt(Varml))
q[5] = pnorm(945 ,Meanml, sqrt(Varml)) - pnorm(895 ,Meanml, sqrt(Varml))
q[6] = pnorm(945 ,Meanml, sqrt(Varml), FALSE)
q
```
We can use the ML estimates in the GLR test statistic.
M = 6, and ntheta = 2, since a Normal distribution has 2 unknown variables, so use chisq(3) for the lower bound and chisq(5) for the upper bound.
```{r}
#create test statistic
c <-  -2*log( prod(q^Obsv) / prod( (Obsv/100)^Obsv ))
c

pchisq(c, 3, lower.tail=FALSE)
pchisq(c, 5, lower.tail=FALSE)
```
Both are greater than 0.05, so the test fails at the 0.05 significance level, and hence we have insufficient grounds to reject H0.



